---
title: "Lab 9 - Programming in R - Regular Expressions"
author: Berent Aldikacti
date: "12/03/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r style, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library("BiocStyle")
library("knitr")
library("rmarkdown")
library(stringr)
library(tidyverse)
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE,
               cache = FALSE, fig.width = 5, fig.height = 5)

```


```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```
```{r}
str_view(x, ".a.") # dot is used to represent any character
```
```{r}
# To escape the regular behaviour of "." we use a escape "\\".  
# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
```

```{r}
# And this tells R to look for an explicit .
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c") # This tells to look for an actual dot instead of the regular behaviour of dot
```
```{r}
x <- "a\\b"
writeLines(x)
str_view(x, "\\\\")
```

## Exercise 14.3.1.1

1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

Backslash is used as a escape in regex and it is also a string.  

"\" : This will escape the next R string character
"\\" : This will resolve to literal \ in regex which will escape the next character regex. 
"\\\" : First two will resolve to literal \ in regex and the last \ will escape a character. In regex, this will escape an escaped character


# 14.3.2 Anchors

* ^ to match the start of the string.
* $ to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "a$")
```

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
str_view(x, "^apple$")
```

```{r}
str_view(x, "\bapple\b")
```


## 14.3.2.1 Exercises

1. How would you match the literal string "$^$"?
```{r}
x <- "$^$"
str_view(x,"^\\$\\^\\$$")
```

# 14.3.3 Character classes and alternatives


* \d : matches any digit.
* \s : matches any whitespace (e.g. space, tab, newline).
* [abc] : matches a, b, or c.
* [^abc] : matches anything except a, b, or c.

```{r}
# Look for a literal character that normally has special meaning in a regex
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
str_view(x, ".[*]c")
str_view(x, "a[ ]")
```

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

## 14.3.3.1 Exercises

1. Create regular expressions to find all words that:
  * Start with a vowel.
  * That only contain consonants. (Hint: thinking about matching “not”-vowels.)
  * End with ed, but not with eed.
  * End with ing or ise.

```{r}
str_subset(stringr::words, "^[a|e|i|o|u]")
```

```{r}
str_subset(stringr::words, "[a|e|i|o|u]", negate = TRUE)
```

```{r}
str_subset(stringr::words, "[^e]ed$")
```

# 14.3.4 Repetition


* ?: 0 or 1
* +: 1 or more
* *: 0 or more


```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```

You can also specify the number of matches precisely:

* {n}: exactly n
* {n,}: n or more
* {,m}: at most m
* {n,m}: between n and m

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

By default these matches are “greedy”: they will match the longest string possible. You can make them “lazy”, matching the shortest string possible by putting a ? after them. This is an advanced feature of regular expressions, but it’s useful to know that it exists:
```{r}
str_view(x, 'C{2,3}?')
```
```{r}
str_view(x, 'C[LX]+?')
```

## 14.3.4.1 Exercises

1. Describe the equivalents of ?, +, * in {m,n} form.

? = {0,1}
+ = {1,}
* = {0,}

# 14.3.5 Grouping and backreferences

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

## 14.3.5.1 Exercises

1. Describe, in words, what these expressions will match:

(.)\1\1 = The same character appearing three times in a row. E.g. "aaa"
"(.)(.)\\2\\1" = A pair of characters followed by the same pair of characters in reversed order. E.g. "abba".
(..)\1 = Any two characters repeated. E.g. "a1a1".
"(.).\\1.\\1" = A character followed by any character, the original character, any other character, the original character again. E.g. "abaca", "b8b.b".
"(.)(.)(.).*\\3\\2\\1" = Three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. E.g. "abcsgasgddsadgsdgcba" or "abccba" or "abc1cba".

# 14.4 Tools

## 14.4.1 Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```


```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```


```{r}
str_count("abababa", "aba")
str_view_all("abababa", "aba")
```

## 14.4.1.1 Exercises

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.

Find all words that start or end with x.
```{r}
str_subset(words, "^x|x$")
```

Find all words that start with a vowel and end with a consonant.
```{r}
str_subset(words, "^[aeiou]|[^aeiou]$")
```

Are there any words that contain at least one of each different vowel?


```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```

# 14.4.2 Extract matches

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
str_extract_all(more, colour_match)
```

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

## 14.4.2.1 Exercises

1. In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)
```

# 14.4.3 Grouped matches

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
has_noun %>% 
  str_match(noun)
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

## 14.4.3.1 Exercises

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```

# 14.4.4 Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

Instead of replacing with a fixed string you can use backreferences to insert components of the match. In the following code, I flip the order of the second and third words.
```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

## 14.4.4.1 Exercises

1. Replace all forward slashes in a string with backslashes.

```{r}
x <- "///\\]/]/]/]/"
x %>% str_replace_all("/","\\\\")
```

# 14.4.5 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```
```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

## 14.4.5.1 Exercises

1. Split up a string like “apples, pears, and bananas” into individual components.

```{r}
x <- "apples, pears, and bananas"
str_split(x, boundary("word"))[[1]]
```
































